Cross-scale integration of knowledge for predicting species ranges: a metamodeling framework

### 2015-08-07 tex/appendix_s1.tex

\usepackage[dvipsnames]{xcolor}
\newcommand{\rev}[1]{{\color{RoyalBlue}#1}}

@@

%==================
% FIGURE 1
\begin{figure}[t]
\input{diagram}
\caption{The parameters of a correlative metamodel model (left column) are conditioned on the predictions of a mechanistic sub-model (right column).
	The metamodel ($\theta_M$) operates at a single scale and uses occurrence data (\(X_M\)) and explanatory variables ($D_M$) to produce a naive (i.e., not conditioned on sub-models) prediction $\psi_N$.
	The mechanistic sub-model \(\theta_S\) includes data about the response (\(X_S\)) of lower-level behaviours of the system to explanatory variables ($D_S$). 
	The models are integrated by calibrating $\theta_M$ to data ($X_M, D_M$) as well as the output of the sub-model ($\psi_S$). 
	This is possible because predictions from the sub-model ($\psi_S$) emerge at the scale of the metamodel via a scaling function \(g(\theta_S, D_S)\).
	This prediction incorporates multiple sources of information coming from several calibration datasets (i.e., $X_M, D_M, X_S, $ and $D_S$) as well as from multiple types of models (i.e., $\theta_M$ and $\theta_S$).}
\label{fig:diagram}
\end{figure}
%==================

@@

For the mechanistic submodel, we computed the predicted probability of presence as a function of precipitation, using the experimental data and the theoretical prediction that the species will be present when the population growth rate is greater than 0.
For each of the five experimental treatments, we computed the probability of presence as the integral of the normal density from 0 to infinity, where the mean of the normal distribution was equal to the average population growth rate at each precipitation treatment and the standard deviation was the corresponding standard error for each treatment.
We then fit these data to the same model, priors, and sample sizes as those from the previous step.
Fitting the metamodel is simply a matter of repeating the exact procedure for fitting the naive model, but using informative priors generated from the naive model step.
Because the sub-model considered only precipitation, and because the response was highly correlated to precipitation, we expected the submodel to produce highly precise estimates of the parameters for the effect of precipitation on the probability of presence.
However, the sub-model was quite simplistic, and thus likely over-estimates its precision when considering the species range wholistically (as is done with the metamodel).
Thus, we applied a prior weight of 0.05 to the sub-model.
This allowed the sub-model to inform the integrated model without dominating the results.

@@

\begin{figure}[t]
\includegraphics[width=4.5in]{figs/plotmap.png}
\caption{Map of permanent forest plot locations}
\label{fig:plotmap}
\end{figure}

We divided the dataset into calibration and validation subsets, with 1/3 of the data reserved for validation and the remaining 2/3 used for calibration.
As in the previous example, we modeled the probability of presence as a function of climate.
We considered seven climate variables: the number of degree days (ddeg), minimum temperature (min\_temp), potential evapotranspiration (PET), annual precipitation (an\_prcp), summer precipitation (sum\_prcp), winter precipitation (win\_prcp), and the ratio of annual precipitation to potential evapotranspiration (pToPET).
See \citet{Morin2009} for a complete description of these variables.
For projection, we used predictions for \rev{2100} from the HadCM3 GCM \citep{Pope2000} driven by the A2 emission scenario \citep{Nakicenovic2000}, and used the parameters of the models to forecast suitability into the future.
Exploratory analysis revealed that there were collinearities within the predictor variables.
Thus, we only included three variables in the model: ddeg, an\_prcp, and pToPET.
To select the form of the naive model (and thus the metamodel), we used stepwise regression with BIC as the evaluation criteria.
The search scope included all models with linear, quadratic, and cubic terms for all three variables.
\rev{Thus, the naive model (which also defined the scope of the metamodel), can be formulated as a simple Binomial GLM, as in Example 1:
\begin{equation}
\psi_N =\text{logit}^{-1}\left( \theta_M D_M \right)
\end{equation}
where \(\psi_N\) is the probability of presence in the naive model, \(\theta_M\) is the parameter vector of the metamodel (presently unconstrained by any additional information), and \(D_M\) is the macroclimate covariate matrix at the metamodel scale.
The parameter vector is estimated following equation 2 in the main text:
\begin{equation}
p\left (\theta_M \mid X_M,D_M \right ) \propto 
p \left(X_M \mid \theta_M, D_M \right)
p \left(\theta_M \right)
\end{equation}

Because Phenofit provided predictions in the form of a probability of presence (rather than directly in terms of the metamodel parameters as in the previous example), it was necessary to relate these predictions to the metamodel.
We opted to do this via simulation because it is a natural method to use in an MCMC scheme.
We thus treated the Phenofit predictions (\(\psi_P\)) as an observed result of a true underlying set of occurrences (\(X_P\)).
Because these occurrences were unobserved (and in the case of the future distribution, unobservable), it was necessary to infer them using MCMC.
For each MCMC iteration, we generated a simulated occurrence dataset \(\hat{X}_P\) by drawing from a Binomial distribution with probabilities equal to \(\psi_P\).
In order to give both the occurrence dataset and Phenofit predictions equal weight, we drew the same number of samples as were present in the occurrence dataset.
The metamodel was then fit to these simulated data (with conditioning on the naive model as well):
\begin{equation} 
\overbrace{p(\theta_M \mid X_M, D_M, X_P, \psi_P)}^\text{integrated posterior}
\propto
\overbrace{p\left (X_P \mid \theta_M,D_M,\psi_P \right )}^{\substack{\text{new information} \\ \text{from Phenofit}}}
\overbrace{p \left(X_M \mid \theta_M, D_M \right) p \left(\theta_M \right)}^{\substack{\text{naive metamodel} \\ \text{posterior}}}
\overbrace{p \left(\psi_P \right)}^{\substack{\text{prior for} \\ \text{Phenofit}}}	
\end{equation}
This procedure, by generating a random dataset with each iteration, incorporated the variance in the Phenofit predictions.
Finally, because we were interested in two separate questions with respect to Phenofit integration (one regarding how integration affects uncertainty in our present estimate of the species distribution and a one regarding the effect of integration on projections into the future), we performed two separate integrations with the Phenofit information.
To address the first question, we used the Phenofit predictions for the present as \(\psi_P\) and present climate as \(D_M\), producing the Integrated-Present model.
To address the second, we used the future predictions from Phenofit and future climate, resulting in the Integrated-Future model.
In both cases, the naive model was a GLM relating present occurrences to present climate (and thus the prior response to the environment was static, similarly to how one might use the same conventional SDM to both model present distribution and project into the future).
Convergence for all models was assessed by comparing the results of multiple chains with overdispersed starting points.
Based on the results of these initial runs, we thinned all posterior samples by recording only every 50th sample, and we obtained 25000 posterior samples for each model following a burn-in period of 20000 samples.}

\section*{Implementation}
\rev{Here, we provide a brief guide on implementing our framework. We focus on our two examples, providing a brief outline of the steps to reproduce the analyses using the code and data provided in Appendix S2.
In the section for each example, we provide some references to software packages that may be of assistance to beginning users.
As a full discussion of the practicalities of Bayesian modelling is beyond the scope of this paper, we direct the reader to some of the many excellent resources for Bayesian and hierarchical modelling in ecology \citep[e.g., we suggest][]{Bolker2007, Royle2009, Link2010}.
We also provide some additional information on applying model weights.
Our intent is that, with this guide and the accompanying code, users familiar with R and with some experience with Bayesian modelling will be able to implement our method.}

\subsection*{Example 1}
\rev{
Example 1 is implemented fully using off-the shelf statistical software.
Data preparation, exploration, and visualization is all performed in R \citep{R}.
Parameterization of all models is performed using JAGS, which is a standalone implementation of a Metropolis-Hastings and Gibbs sampler, as well as RJAGS, the R interface to the JAGS program \citep{RJAGS}.
Within the {\tt example\_1} directory in Appendix S2, we provide 6 numbered scripts.
These are designed to be run from the command line using the RScript program that ships with most standard R installations (e.g., {\tt Rscript 1-m1.r} to run the first script).
However, we encourage users to review the code rather than simply running the examples from the command line.
Users of the R console (or the RStudio environment) will be able to run the code one line at a time, provided the working directory is set to the {\tt example\_1} directory beforehand.

There are three main steps to running the analysis, corresponding to the first three numbered scripts.
The first (file {\tt 1-m1.r}) generates the simulated presence-absence dataset, loads a JAGS model representing the metamodel (from the file {\tt ex1\_metamodel.jags}), and runs JAGS on the model to produce posterior samples for the naive model.
Note that, because the naive model and the metamodel are at the same scale and differ only in whether they include additional information from the sub-model, the JAGS model file is the same for both.
All data and results from this step are saved to the {\tt dat/ex1\_m1.rdata} file.
Step two ({\tt 2-m2.r}) creates the dataset for the sub-model, loads the JAGS model ({\tt ex1\_model2.jags}), and runs the model, saving all results to {\tt dat/ex1\_m2.rdata}.
Finally, the third step ({\tt 3-mm.r}) simply reads the estimates from the sub-model, incorporates them into the information passed to JAGS for estimating the metamodel, and runs JAGS to obtain the integrated posterior samples and save them to {\tt dat/ex1\_mm.rdata}.
The posterior samples are saved as MCMC objects from package coda; users should refer to the documentation for that package for information on using these objects.
The remaining three scripts contain code for reproducing the figures contained in the main manuscript.
Users may find them useful for guidance on working with posterior samples generated with JAGS.
}

\subsection*{Example 2}
\rev{
For performance reasons, we have implemented Example 2 as a standalone C++ program, designed to be run multi-threaded in high-performance computing applications.
Due to the size of the dataset and the increased complexity of the model (relative to Example 1), the computational load is considerably higher.
The procedure to compile and run the model is also somewhat more complex.
Therefore, to facilitate exploration without requiring users to actually build and run the model, we have included the posterior samples we obtained with Appendix S2.
This will allow users to build the figures and explore the results of the naive and integrated models.
For the portions that can be run in R, as in Example 1, we have included numbered scripts that can be run with {\tt Rscript}.
There are three scripts included in the {\tt example\_2/r/} directory, numbered 1, 3, and 5 (steps 2 and 4 are running the C++ program to obtain samples for the naive and integrated models, respectively).
Step 1 ({\tt 1\_prepare\_naive\_data.r}) prepares the dataset for the naive model to be used with the sampler, resulting in the files {\tt dat/mcmc/naiveData.csv} (for use with the C++ program) and {\tt dat/mcmc/naiveData.rds} (for use with R, can be loaded using the {\tt readRDS} function).
Step 2 (see following paragraph) creates the file {\tt results/mcmc/naivePosterior.csv}, which contains the posterior samples for the naive model and which we have included for users that do not wish to build and run the C++ program themselves.
Step 3 reads the results of the naive model estimation and uses it to prepare the files needed to estimate the integrated model.
As with step 2, step 4 depends on the C++ program and produces posterior samples for both the Integrated-Present and Integrated-Future models (files {\tt integratedPresent.csv} and {\tt integratedFuture.csv}, respectively, in the {\tt results/mcmc/} directory).
Finally, step 5 reads all of the CSV files containing posterior samples and processes them into MCMC objects (saved in {\tt results/posteriors.rdata}) that can be easily manipulated in R with the coda package.
We also include code for reproducing the figures contained in the manuscript (all scripts starting with {\tt fig}), and a brief interactive script for performing model evaluation ({\tt evaluation.r})

For users wishing to build and run the C++ program to estimate the posterior distributions, some knowledge of C++ compilers will be required.
A familiarity with GNU Make will also be of assistance, because the build procedure is fully automated and documented within the included makefile.
All C++ code is contained within the {\tt src} subdirectory and should be fully compatible with any platform capable of compiling and running standard C++89 programs.
The only external dependency is the GNU Scientific Library (freely available at http://www.gnu.org/software/gsl/), an open-source C library which we have used for random number generation and other mathematical operations.
Additionally, for high-perfomance computing applications, we have included compiler commands for using OpenMP.
When enabled, this will significantly improve the speed at which the likelihood is computed, and thus reducing the runtime of the model.
All compiler commands, as well as the commands used to run the model and generate posterior samples, are included in the makefile.
Additionally, once the program is compiled, help with program options can be obtained using the {\tt -h} switch (i.e., {\tt integrated\_model -h} at the command line).
}

\subsection*{Model weights}
\rev{Model weights, expressed mathematically as the prior probability of each individual sub-model, allow the user to express prior belief that a particular sub-model is effective at describing the range of a species.
This is particularly useful, as in Example 1, when a sub-model is highly statistically precise but, due to limited scope, is not effective alone at describing the range of a species.
The hypothetical experiment in Example 1 included only 5 treatments in a single variable, and thus misses many important processes that generate species ranges.
Selecting the specific weighting to use will, of necessity, be somewhat subjective, as it is often difficult to express prior belief of this sort in precise mathematical terms.
As a start, users may consider weighting models based on sample size or number of independent treatments, as we have done in example 1, where the weight for the sub-model of 0.05 was chosen based on the 5 precipitation levels relative to 100 independent data points in the correlative dataset.
Implementation of the weights within the model code itself will depend on the implementation.
In the case of Example 1, the weight was applied by decreasing the precision of all terms that were informed by the submodel (i.e., making the prior that resulted from the sub-model less informative).
}

>>>>

@@

%==================
% FIGURE 1
\begin{figure}[p!]
\input{diagram}
\caption{The parameters of a correlative metamodel model (left column) are conditioned on the predictions of a mechanistic sub-model (right column).
	The metamodel ($\theta_M$) operates at a single scale and uses occurrence data (\(X_M\)) and explanatory variables ($D_M$) to produce a naive (i.e., not conditioned on sub-models) prediction $\psi_N$.
	The mechanistic sub-model \(\theta_S\) includes data about the response (\(X_S\)) of lower-level behaviours of the system to explanatory variables ($D_S$). 
	The models are integrated by calibrating $\theta_M$ to data ($X_M, D_M$) as well as the output of the sub-model ($\psi_S$). 
	This is possible because predictions from the sub-model ($\psi_S$) emerge at the scale of the metamodel via a scaling function \(g(\theta_S, D_S)\).
	This prediction incorporates multiple sources of information coming from several calibration datasets (i.e., $X_M, D_M, X_S, $ and $D_S$) as well as from multiple types of models (i.e., $\theta_M$ and $\theta_S$).}
\label{fig:diagram}
\end{figure}
%==================

@@

For the mechanistic submodel, we computed the predicted probability of presence as a function of precipitation, using the experimental data and the theoretical prediction that the species will be present when the population growth rate is greater than 0.
For each of the five experimental treatments, we computed the probability of presence as the integral of the normal density from 0 to infinity, where the mean of the normal distribution was equal to the average population growth rate at each precipitation treatment and the standard deviation was the corresponding standard error for each treatment.
We then fit these data to the same model, priors, and sample sizes as those from the previous step.
Fitting the metamodel is simply a matter of repeating the exact procedure for fitting the naive model, but using informative priors generated from the naive model step.
Because the sub-model considered only precipitation, and because the response was highly correlated to precipitation, we expected the submodel to produce highly precise estimates of the parameters for the effect of precipitation on the probability of presence.
However, the sub-model was quite simplistic, and thus likely over-estimates its precision when considering the species range wholistically (as is done with the metamodel).
Thus, we applied a prior weight of 0.05 to the sub-model.
This allowed the sub-model to inform the integrated model without dominating the results.
For temperature-related parameters, the posterior distributions for the naive and integrated models were indistinguishable, whereas parameters related to precipitation were substantially revised, including large reductions in uncertainty for the integrated model (Fig. \ref{fig:ex1params}).

\begin{figure}
\includegraphics[width=6.5in]{figs/ex1_params.png}
\caption{Relative posterior density of all example one parameters for the naive model, mechanistic sub-model, and the integrated model.
Parameters related to temperature ($b_1$ and $b_2$) were not informed with additional data, and showed little change from the naive model.
The intercept ($b_0$ and parameters related to precipitation ($b_3, b_4$) were informed by integration and showed substantial revision between the naive and integrated models.
Boxplots show the median, interquartile range, and the range.}
\label{fig:ex1params}
\end{figure}

@@

\begin{figure}
\includegraphics[width=4.5in]{figs/plotmap.png}
\caption{Map of permanent forest plot locations}
\label{fig:plotmap}
\end{figure}

We divided the dataset into calibration and validation subsets, with 1/3 of the data reserved for validation and the remaining 2/3 used for calibration.
As in the previous example, we modeled the probability of presence as a function of climate.
We considered seven climate variables: the number of degree days (ddeg), minimum temperature (min\_temp), potential evapotranspiration (PET), annual precipitation (an\_prcp), summer precipitation (sum\_prcp), winter precipitation (win\_prcp), and the ratio of annual precipitation to potential evapotranspiration (pToPET).
See \citet{Morin2009} for a complete description of these variables.
For projection, we used predictions for 2100 from the HadCM3 GCM \citep{Pope2000} driven by the A2 emission scenario \citep{Nakicenovic2000}, and used the parameters of the models to forecast suitability into the future.
Exploratory analysis revealed that there were collinearities within the predictor variables.
Thus, we only included three variables in the model: ddeg, an\_prcp, and pToPET.
To select the form of the naive model (and thus the metamodel), we used stepwise regression with BIC as the evaluation criteria.
The search scope included all models with linear, quadratic, and cubic terms for all three variables.
Thus, the naive model (which also defined the scope of the metamodel), can be formulated as a simple Binomial GLM, as in Example 1:
\begin{equation}
\psi_N =\text{logit}^{-1}\left( \theta_M D_M \right)
\end{equation}
where \(\psi_N\) is the probability of presence in the naive model, \(\theta_M\) is the parameter vector of the metamodel (presently unconstrained by any additional information), and \(D_M\) is the macroclimate covariate matrix at the metamodel scale.
The parameter vector is estimated following equation 2 in the main text:
\begin{equation}
p\left (\theta_M \mid X_M,D_M \right ) \propto 
p \left(X_M \mid \theta_M, D_M \right)
p \left(\theta_M \right)
\end{equation}

Because Phenofit provided predictions in the form of a probability of presence (rather than directly in terms of the metamodel parameters as in the previous example), it was necessary to relate these predictions to the metamodel.
We opted to do this via simulation because it is a natural method to use in an MCMC scheme.
We thus treated the Phenofit predictions (\(\psi_P\)) as an observed result of a true underlying set of occurrences (\(X_P\)).
Because these occurrences were unobserved (and in the case of the future distribution, unobservable), it was necessary to infer them using MCMC.
For each MCMC iteration, we generated a simulated occurrence dataset \(\hat{X}_P\) by drawing from a Binomial distribution with probabilities equal to \(\psi_P\).
In order to give both the occurrence dataset and Phenofit predictions equal weight, we drew the same number of samples as were present in the occurrence dataset.
The metamodel was then fit to these simulated data (with conditioning on the naive model as well):
\begin{equation} 
\overbrace{p(\theta_M \mid X_M, D_M, X_P, \psi_P)}^\text{integrated posterior}
\propto
\overbrace{p\left (X_P \mid \theta_M,D_M,\psi_P \right )}^{\substack{\text{new information} \\ \text{from Phenofit}}}
\overbrace{p \left(X_M \mid \theta_M, D_M \right) p \left(\theta_M \right)}^{\substack{\text{naive metamodel} \\ \text{posterior}}}
\overbrace{p \left(\psi_P \right)}^{\substack{\text{prior for} \\ \text{Phenofit}}}	
\end{equation}
This procedure, by generating a random dataset with each iteration, incorporated the variance in the Phenofit predictions.
Finally, because we were interested in two separate questions with respect to Phenofit integration (one regarding how integration affects uncertainty in our present estimate of the species distribution and a one regarding the effect of integration on projections into the future), we performed two separate integrations with the Phenofit information.
To address the first question, we used the Phenofit predictions for the present as \(\psi_P\) and present climate as \(D_M\), producing the Integrated-Present model.
To address the second, we used the future predictions from Phenofit and future climate, resulting in the Integrated-Future model.
In both cases, the naive model was a GLM relating present occurrences to present climate (and thus the prior response to the environment was static, similarly to how one might use the same conventional SDM to both model present distribution and project into the future).
Convergence for all models was assessed by comparing the results of multiple chains with overdispersed starting points.
Based on the results of these initial runs, we thinned all posterior samples by recording only every 50th sample, and we obtained 25000 posterior samples for each model following a burn-in period of 20000 samples.
Finally, we evaluated the Naive and Integrated-Present models by computing the area under the receiver operating curve (AUC).
AUC measures the ability of the model to discriminate presences and absences, where a value of 1 indicates perfect discrimination and 0.5 is random discrimination.
Because we obtained joint posterior distributions for all models, it is not possible to evaluate the performance of a `best' model; rather, we computed AUC for every posterior sample of both models, resulting in posterior distributions of AUC (Fig. \ref{fig:auc}).
Median AUC values were 0.802 and 0.797 for the Naive and Integrated-Present models, respectively, and the overlap between the two distributions suggests little difference in the performance of each model.

\begin{figure}
\includegraphics[width=5.5in]{figs/ex2_auc.png}
\caption{Posterior distributions of the area under the receiver operating curve (AUC) for the Naive and Integrated-Present models. For comparison, the y-values have been scaled by the maximum densities. Boxplots in the upper panel show the median, interquartile range and \(1.5 \times\) the interquartile range, along with more extreme values plotted individually.}
\label{fig:auc}
\end{figure}

\section*{Implementation}
Here, we provide a brief guide on implementing our framework. We focus on our two examples, providing a brief outline of the steps to reproduce the analyses using the code and data provided in Appendix S2.
In the section for each example, we provide some references to software packages that may be of assistance to beginning users.
As a full discussion of the practicalities of Bayesian modelling is beyond the scope of this paper, we direct the reader to some of the many excellent resources for Bayesian and hierarchical modelling in ecology \citep[e.g., we suggest][]{Bolker2007, Royle2009, Link2010}.
We also provide some additional information on applying model weights.
Our intent is that, with this guide and the accompanying code, users familiar with R and with some experience with Bayesian modelling will be able to implement our method.

\subsection*{Example 1}
Example 1 is implemented fully using off-the shelf statistical software.
Data preparation, exploration, and visualization is all performed in R \citep{R}.
Parameterization of all models is performed using JAGS, which is a standalone implementation of a Metropolis-Hastings and Gibbs sampler, as well as RJAGS, the R interface to the JAGS program \citep{RJAGS}.
Within the {\tt example\_1} directory in Appendix S2, we provide 6 numbered scripts.
These are designed to be run from the command line using the RScript program that ships with most standard R installations (e.g., {\tt Rscript 1-m1.r} to run the first script).
However, we encourage users to review the code rather than simply running the examples from the command line.
Users of the R console (or the RStudio environment) will be able to run the code one line at a time, provided the working directory is set to the {\tt example\_1} directory beforehand.

There are three main steps to running the analysis, corresponding to the first three numbered scripts.
The first (file {\tt 1-m1.r}) generates the simulated presence-absence dataset, loads a JAGS model representing the metamodel (from the file {\tt ex1\_metamodel.jags}), and runs JAGS on the model to produce posterior samples for the naive model.
Note that, because the naive model and the metamodel are at the same scale and differ only in whether they include additional information from the sub-model, the JAGS model file is the same for both.
All data and results from this step are saved to the {\tt dat/ex1\_m1.rdata} file.
Step two ({\tt 2-m2.r}) creates the dataset for the sub-model, loads the JAGS model ({\tt ex1\_model2.jags}), and runs the model, saving all results to {\tt dat/ex1\_m2.rdata}.
Finally, the third step ({\tt 3-mm.r}) simply reads the estimates from the sub-model, incorporates them into the information passed to JAGS for estimating the metamodel, and runs JAGS to obtain the integrated posterior samples and save them to {\tt dat/ex1\_mm.rdata}.
The posterior samples are saved as MCMC objects from package coda; users should refer to the documentation for that package for information on using these objects.
The remaining three scripts contain code for reproducing the figures contained in the main manuscript.
Users may find them useful for guidance on working with posterior samples generated with JAGS.

\subsection*{Example 2}
For performance reasons, we have implemented Example 2 as a standalone C++ program, designed to be run multi-threaded in high-performance computing applications.
Due to the size of the dataset and the increased complexity of the model (relative to Example 1), the computational load is considerably higher.
The procedure to compile and run the model is also somewhat more complex.
Therefore, to facilitate exploration without requiring users to actually build and run the model, we have included the posterior samples we obtained with Appendix S2.
This will allow users to build the figures and explore the results of the naive and integrated models.
For the portions that can be run in R, as in Example 1, we have included numbered scripts that can be run with {\tt Rscript}.
There are three scripts included in the {\tt example\_2/r/} directory, numbered 1, 3, and 5 (steps 2 and 4 are running the C++ program to obtain samples for the naive and integrated models, respectively).
Step 1 ({\tt 1\_prepare\_naive\_data.r}) prepares the dataset for the naive model to be used with the sampler, resulting in the files {\tt dat/mcmc/naiveData.csv} (for use with the C++ program) and {\tt dat/mcmc/naiveData.rds} (for use with R, can be loaded using the {\tt readRDS} function).
Step 2 (see following paragraph) creates the file {\tt results/mcmc/naivePosterior.csv}, which contains the posterior samples for the naive model and which we have included for users that do not wish to build and run the C++ program themselves.
Step 3 reads the results of the naive model estimation and uses it to prepare the files needed to estimate the integrated model.
As with step 2, step 4 depends on the C++ program and produces posterior samples for both the Integrated-Present and Integrated-Future models (files {\tt integratedPresent.csv} and {\tt integratedFuture.csv}, respectively, in the {\tt results/mcmc/} directory).
Finally, step 5 reads all of the CSV files containing posterior samples and processes them into MCMC objects (saved in {\tt results/posteriors.rdata}) that can be easily manipulated in R with the coda package.
We also include code for reproducing the figures contained in the manuscript (all scripts starting with {\tt fig}), and a brief interactive script for performing model evaluation ({\tt evaluation.r})

For users wishing to build and run the C++ program to estimate the posterior distributions, some knowledge of C++ compilers will be required.
A familiarity with GNU Make will also be of assistance, because the build procedure is fully automated and documented within the included makefile.
All C++ code is contained within the {\tt src} subdirectory and should be fully compatible with any platform capable of compiling and running standard C++98 programs.
The only external dependency is the GNU Scientific Library (freely available at http://www.gnu.org/software/gsl/), an open-source C library which we have used for random number generation and other mathematical operations.
Additionally, for high-perfomance computing applications, we have included compiler commands for using OpenMP.
When enabled, this will significantly improve the speed at which the likelihood is computed, and thus reducing the runtime of the model.
All compiler commands, as well as the commands used to run the model and generate posterior samples, are included in the makefile.
Additionally, once the program is compiled, help with program options can be obtained using the {\tt -h} switch (i.e., {\tt integrated\_model -h} at the command line).

\subsection*{Model weights}
Model weights, expressed mathematically as the prior probability of each individual sub-model, allow the user to express prior belief that a particular sub-model is effective at describing the range of a species.
This is particularly useful, as in Example 1, when a sub-model is highly statistically precise but, due to limited scope, is not effective alone at describing the range of a species.
The hypothetical experiment in Example 1 included only 5 treatments in a single variable, and thus misses many important processes that generate species ranges.
Selecting the specific weighting to use will, of necessity, be somewhat subjective, as it is often difficult to express prior belief of this sort in precise mathematical terms.
As a start, users may consider weighting models based on sample size or number of independent treatments, as we have done in example 1, where the weight for the sub-model of 0.05 was chosen based on the 5 precipitation levels relative to 100 independent data points in the correlative dataset.
Implementation of the weights within the model code itself will depend on the implementation.
In the case of Example 1, the weight was applied by decreasing the precision of all terms that were informed by the submodel (i.e., making the prior that resulted from the sub-model less informative).
